---
title: "p8105_hw5_zn2220"
author: "Ziang Niu"
date: "2025-10-31"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)

set.seed(235)
```

# Problem 1

## Write a function that, for a fixed group size, randomly draws “birthdays” for each person; checks whether there are duplicate birthdays in the group; and returns TRUE or FALSE based on the result.

Write the function.

```{r}
duplicate_birthday <- function(n) {
birthdays <- sample(1:365, size = n, replace = TRUE)
any(duplicated(birthdays))
}
```

Check returns.

```{r}
duplicate_birthday(2)
duplicate_birthday(23)
duplicate_birthday(50)
```

## Next, run this function 10000 times for each group size between 2 and 50. For each group size, compute the probability that at least two people in the group will share a birthday by averaging across the 10000 simulation runs. Make a plot showing the probability as a function of group size, and comment on your results.

Create a tibble.

```{r}
n <- 2:50
repeats <- 10000

prob <- numeric(length(n))

for (i in seq_along(n)) {
  group_size <- n[i]
  prob[i] <- mean(replicate(repeats, duplicate_birthday(group_size)))
}

prob_list <- tibble(n = n, probability = prob)

```

draw a plot.

```{r}
ggplot(prob_list, aes(x = n, y = prob)) +
geom_line() +
geom_point() +
scale_x_continuous(breaks = seq(0, 50, by = 5)) +
labs(title = "Probability that at least two people share a birthday",
x = "Group size (n)",
y = "Probability") +
theme_minimal(base_size = 12)
```

As n increases, the probability of shared birthdays rises quickly; at around n = 23, it’s already about 0.5.

